---
title: "Visualization_group_by"
author: "Amadeia Rector"
date: "10/9/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))

```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())


```
The theme sets how it will look - black and white.
Load in weather data set
Doing a call to a package - creating a month variable using lubridate package
-rounded date to nearest month!
-because we might want to look at weather within a month

### Start a grouping
```{r}
weather_df %>% 
  group_by(name)
```

```{r}
weather_df %>% 
  group_by(month)
```
```{r}
weather_df %>% 
  group_by(name, month)
```
It's creating layers.
12X3= 36 month name groups
you can ungroup by same thing too.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup()
```

#### Summarize  
looking at unique elements
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n())
```
number is the name of your summary, and it's looking at the number which is n()
The result is just the dataframe and the summary I computed inside

#### Count first
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(number = n())
```
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(number = n())
```
got 3x number days because for all three weather stations

```{r}
weather_df %>% 
 count(month)
```
count does same as summarize, doesn't create a dataframe

## Grouped summaries
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            num_na = sum(is.na(tmax)),
            mean_tmax = mean(tmax)) %>% 
  View()
```
na.rm =TRUE removes missing values
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            median_prcip = median(prcp, na.rm = TRUE),
            mean_tmax = mean(tmax,na.rm = TRUE)) %>%
  View()
```
Could use spread by name or month to make it human-readable!
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```
computed a mean,  and created a single number value for the group

## window functions can give rank order
use rank
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(tmax_rank = min_rank(tmax)) %>% 
  View()
  mutate()
```

Offsetting data so you can look at difference
use lag ranks
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(lag_max =lag(tmax)) %>% View
```


